A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-08-29 reports/Draft.Rnw
worked on the draft a bit

\begin{abstract}
This paper implements the hedonic price method to estimate the impact of traffic noise on properties within the St.\ Paul, Minnesota metropolitan real-estate market over the time period from 2005 - 2010. We use Locally Weighted Regression (LWR) estimation techniques in order to allow the hedonic function to vary over time and space. Significant evidence suggests spatial and temporal non-stationarity in the hedonic price function within our study area. In contrast to previous research, however, we find little evidence that the impact of traffic noise exhibits a significant non-linear relationship with the decibel level of the traffic noise, nor does the impact seem to vary over space. We find significant differences in the impact of traffic noise from the beginning to the end of our sample. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September, 2008, and an average of 0.37 percent after September, 2008. While significantly different, both of these estimates are still within the typical range of previous estimates in the literature.
\end{abstract}

@@

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression.  Such work has provided important insight on the nature of this relationship.  Thus, \citet{Bateman2001} found that the negative eﬀect of traﬃc noise ranged from 0.08\% to 2.22\% per dB with a mean around 0.55\% per dB. In a review of 65 noise evaluation studies \citet{Navrud2002} found an average decrease of 0.64\% per dB. \textbf{We need to briefly review other relevant papers here.}  However, with the exception of a few studies that have focused on air traffic noise exposure \textbf{(we should check on this)}, the relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX).  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.

>>>>

\begin{abstract}
This paper implements the hedonic price method to estimate the impact of traffic noise on properties within the St.\ Paul, Minnesota metropolitan real-estate market over the time period from 2005 - 2010. We use Locally Weighted Regression (LWR) estimation techniques in order to allow the hedonic function to vary over time and space. Significant evidence suggests spatial and temporal non-stationarity in the hedonic price function within our study area. In contrast to previous research, however, we find little evidence that the price semi-elasticity of traffic noise varies with the decibel level of the traffic noise, nor does the impact seem to exhibit spatial non-stationarity. We find significant differences in the impact of traffic noise from the beginning to the end of our sample. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September, 2008, and an average of 0.37 percent after September, 2008. While significantly different, both of these estimates are still within the typical range of previous estimates in the literature.
\end{abstract}

@@

An important first step towards such an analysis is to examine real estate price fluctuations as a function of noise levels using hedonic regression.  Such work has provided important insight on the nature of this relationship.  Thus, \citet{Bateman2001} found that the negative eﬀect of traﬃc noise ranged from 0.08 to 2.22 percent per dB with a mean around 0.55 percent per dB. In a review of 65 noise evaluation studies \citet{Navrud2002} found an average decrease of 0.64 percent per dB. \textbf{We need to briefly review other relevant papers here.}  However, with the exception of a few studies that have focused on air traffic noise exposure \textbf{(see, for instance, \citet{coh})}, the relationship between road traffic noise levels and real estate prices has rarely been explored in the United States, even though the former is perhaps the greatest source of noise in residential neighborhoods (XXX).  One major reason for the thinness of the scholarly literature on this topic in the US lies in the difficulty of modeling noise propagation over the landscape.
